The role of viscosity on drop impact forces

### 2024-08-28 Rebuttal.R2/Response to Reviewer 2.tex

\maketitle

\textit{The authors have made some improvements to the manuscript based on the feedback provided, however some concerns were not fully addressed, as described below. I will use the same number scheme as in the prior report for consistency. I am content with the response for points that are not re-addressed below.}

We thank the referee for carefully reading our manuscript and providing valuable feedback and suggestions. We have reviewed the referee's comments and made changes based on their suggestions. Below, we offer a point-to-point reply to each of the referee's comments and include the changes made in the manuscript. The referee's comments are in italics, and our replies are in plain black. Changes in the manuscript are highlighted in magenta.

\begin{enumerate}
	\item[$\bullet$] \textit{While I appreciate the inclusion of two additional force datasets, no new droplet shape profiles were included. For a paper that has a significant experimental component, I am surprised with the lack of direct visualization of the droplet dynamics. At minimum, I would recommend adding droplet shape profile comparisons for the two new data sets. Furthermore, these visualizations (including the one already included) should include a scale bar. Additionally, I would suggest including some experimental videos as supplementary materials.}\\[1mm]

@@

	\item[$\bullet$] \textit{2. I appreciate the clarification regarding the error bars. However, a proper characterization (and description of the procedure by which they are determined) of the estimated parametric errors is still incomplete. Rather than including additional discussion, the authors reference the supplementary materials of a prior work (Zhang et al. (2022)). In reviewing this work, I cannot identify error assessments on key parameters (such as impact velocity). An error is now included on the droplet size, but there is no discussion of how drop size is determined nor how this error determined. Additionally, the related point regarding horizontal error bars was not addressed. Given that there is no length limit, I would suggest the authors include all details in the present manuscript rather than referring to incomplete discussions in prior work. (Also a minor point: the new horizontal line associated with the force sensor resolution should be described in the caption.)}\\[1mm]

@@

	\item[$\bullet$] \textit{3.}
	\begin{itemize}
		\renewcommand\labelitemi{--}
		\item \textit{In the experimental methods, the authors now describe that the Bond number is fixed at 1. Given the change in fluid properties between different solutions (with fixed droplet size), this cannot be the case. This needs to be made more precise. Furthermore, in the authors' prior work (Zhang et al. (2022)), it is mentioned that the Bo is fixed at 0.5 in simulation for similar experimental parameters. Why the change?}

		\item \textit{By now providing the drop size used, the Oh and Bo now can be used to define the parameters. However, it is not clear that the non-dimensional values are correct. For instance, Figures 2 and 3 both mention experimental conditions where Oh = 0.0025 whereas apparently different droplet sizes were used (2.05 mm in Figure 2, and 2.54 mm in Figure 3). Was a different fluid used between these figures such that the Oh could be held fixed? To clarify all of these issues, and for the ease of the readers, I strongly suggest the authors add the dimensional parameters for all experiments to the captions (as they have now done for Figure 2), an appendix, or make experimental data available in a supplementary data set.}

		\item \textit{Given the significant amount of data overlap in the various figures (e.g. Figure 3(b)), I am more convinced now that the parameters and raw data need to be provided in a supplementary dataset for reproducibility and to facilitate further comparisons.}\\[1mm]

@@

		Drops with Ohnesorge numbers of $0.0025$, $0.0625$, and $0.2$ have diameters of $2.05 \pm 0.13\,\si{\milli\meter}$, $2.52 \pm 0.11\,\si{\milli\meter}$, and $2.54 \pm 0.09\,\si{\milli\meter}$, respectively. We share reviewer's assertion that it would be a great idea to share all the experimental data. We have revised our GitHub repository to include this information. We will make it available using Zenodo.

		\item \textit{The authors mention that their mixtures maintain ``a fairly constant surface tension and density, around 61 mN/m and 1000 kg/m$^3$, respectively.'' Since the values are now clearly specified in the table now, what is the meaning of these particular characteristic values which do in fact vary by around 20\%?}\\[1mm]

@@

		We thank the reviewer for bringing this to our attention and acknowledge that our original statement could be more precise
		To address this concern, we have revised the text to provide a more accurate description of the property variations:

		\S~\red{2.1:}\\
		\oo
		The properties of the impacting drop are controlled using water-glycerol mixtures with viscosities $\eta_d$ varying by almost two orders of magnitude, from $1\,\si{\milli\pascal}\si{\second}$ to $80.2\,\si{\milli\pascal}\si{\second}$. Surface tension is either $72,\si{\milli\newton}/\si{\meter}$ (pure water) or $61,\si{\milli\newton}/\si{\meter}$ (glycerol), while density $\rho_d$ ranges from $1000,\si{\kilo\gram}/\si{\cubic\meter}$ to $1220,\si{\kilo\gram}/\si{\cubic\meter}$, as detailed in table~1 \citep{cheng2008formula, volk2018density, Jha2020}.
		\bb
	\end{itemize}

	\item[$\bullet$] \textit{5. The new title is more appropriate, but it should be mentioned in the title that there are specific restrictions/assumptions on the substrate (i.e. non-wetting), as the results are likely to depend on the surface wettability.}

	We appreciate the reviewer's suggestion regarding the title. While we understand the importance of highlighting specific assumptions, we believe the current title strikes an appropriate balance between conciseness and informativeness. Nevertheless, we would like to address the reviewer's concern by elaborating on our reasoning:

	\begin{enumerate}
		\item \textbf{Generality of the first impact peak:} As demonstrated in the supplementary information of \citet{zhang2022impact}, the first impact peak is independent of surface wettability. 
		\item \textbf{Early disclosure of surface properties:} We explicitly mention the non-wetting nature of the surface in the second line of the abstract. This ensures that readers are immediately aware of this crucial experimental condition without overburdening the title.
	\end{enumerate}

	We acknowledge that the second peak is indeed dependent on surface wettability. However, this specific aspect is thoroughly discussed within the paper, and we believe it does not warrant inclusion in the title.

	\item[$\bullet$] \textit{7. If the experiment is repeatable, for a given drop height, the oscillation phase at which the droplet arrives at the surface should not vary, and thus is unlikely to be captured by the error bars as claimed. Some additional quantification of the non-sphericity of the droplets should be included.}\\[1 mm]

	The reviewer is correct that for a given drop height, the oscillation phase should be consistent if all other factors remain constant. In our experiments, we constrain our analysis to drops with aspect ratios (horizontal to vertical diameter) between 0.96 and 1.05. Given this narrow range, we posit that the impact of these shape variations is negligible compared to the experimental error bars derived from repeated trials under identical nominal conditions. We have modified the text to address this point and provide additional quantification of the non-sphericity of the droplets as suggested by the reviewer:

	\S~\red{2.2:}\\
	\oo
	It is important to note that experimental drops may deviate from perfect sphericity due to air drag post-needle release and potential residual oscillations from detachment. These shape perturbations are more pronounced in cases with low Weber and Ohnesorge numbers. To quantify this non-sphericity, we measure the drop's aspect ratio (horizontal to vertical diameter) immediately before substrate contact. The precise pre-impact drop shape can significantly influence subsequent impact dynamics \citep{thoraval-2013-jfm, yun2017bouncing,Zhang2019}. In our experiments, we constrain our analysis to drops with aspect ratios between 0.96 and 1.05. Given this narrow range, we posit that the impact of these shape variations is negligible compared to the experimental error bars derived from repeated trials under identical nominal conditions.
	\bb

### 2024-08-28 Main/ViscousDropImpact_v7.tex
feat: Update title to highlight non-wetting surface assumption
The title has been updated to more accurately reflect the specific surface wettability assumption used in the study. While the first impact peak is independent of wettability, the second peak is dependent on the non-wetting nature of the substrate. This change ensures the title provides clear information about the experimental conditions to the reader.

Additionally, the manuscript has been updated to better quantify the non-sphericity of the droplets prior to impact, addressing the reviewer's concern about the potential impact of

We leverage the axial symmetry of the drop impact (figure~\ref{fig:schematic}a). This axial symmetry breaks at large $We$ ($\ge 100$ for water drops and even larger Weber number for more viscous drops), owing to destabilization by the surrounding gas after splashing \citep{xu2005drop, Eggers2010, Driscoll2011, riboux2014experiments, Josserand2016, zhang2022impact}. 
To solve the governing equations~\eqref{eqn:continuity}-\eqref{eqn:VoF}, the velocity field $\boldsymbol{v}$ and time $t$ are normalized by the inertio-capillary scales, $V_\gamma = \sqrt{\gamma/\rho_dD_0}$ and $\tau_\gamma = \sqrt{\rho_dD_0^3/\gamma}$, respectively. Furthermore, the pressure is normalized using the capillary pressure scale $p_\gamma = \gamma/D_0$. In such a conceptualization, $Oh$ and $We$ described in \S\ref{sec:intro} uniquely determine the system. 
The Ohnesorge number based on air viscosity $Oh_a = \left(\eta_a/\eta_d\right)Oh$ and air-drop density ratio $\rho_a/\rho_d$ are fixed at $10^{-5}$ and $10^{-3}$, respectively to minimize the influence of the surrounding medium on the impact forces. 
\Rev{Lastly, we keep the Bond number ($Bo$, see~\eqref{Eq:BoDefinition}) fixed at 1 throughout the manuscript. In our system, the relevance of gravity is characterized by the dimensionless Froude number $Fr = V_0^2/gD_0 = We/Bo$ comparing inertia with gravity. Throughout this manuscript, $Fr > 1$ and gravity's role is sub-dominant compared to inertia (for detailed discussion, see appendix~\ref{app:gravity}).}
The substrate is modeled as a no-slip and non-penetrable wall, whereas vanishing stress and pressure are applied at the remaining boundaries to mimic outflow conditions for the surrounding air. The domain boundaries are far enough from the drop not to influence its impact process ($\mathcal{L}_\text{max} \gg D_0$, $\mathcal{L}_\text{max} = 8R$ in the worst case). At $t = 0$, in our simulations, we release \Rev{a spherical drop (see appendix~\ref{app:shape})} whose south pole is $0.05D_0$ away from the substrate and is falling with a velocity $V_0$. The simulations utilize adaptive mesh refinement to finely resolve the velocity, viscous dissipation, and the VoF tracer fields. A minimum grid size $\Delta = D_0/2048$ is used for this study.

@@

\section{Role of drop shape on impact forces}
\label{app:shape}

It is important to note that the drops may not be perfectly spherical and may have residual oscillations as they are released from the needle and arrive at the substrate. These oscillations are expected to be more pronounced for cases with small Weber and Ohnesorge numbers. The exact shape of the drop just before impact can influence the impact dynamics \citep{thoraval-2013-jfm, yun2017bouncing}. However, we suspect that the influence of these shape variations is captured, at least in part, by the error bars in the experimental data, which are derived from repeated trials under the same nominal conditions.

>>>>

We leverage the axial symmetry of the drop impact (figure~\ref{fig:schematic}a). This axial symmetry breaks at large $We$ ($\ge 100$ for water drops and even larger Weber number for more viscous drops), owing to destabilization by the surrounding gas after splashing \citep{xu2005drop, Eggers2010, Driscoll2011, riboux2014experiments, Josserand2016, zhang2022impact}. 
To solve the governing equations~\eqref{eqn:continuity}-\eqref{eqn:VoF}, the velocity field $\boldsymbol{v}$ and time $t$ are normalized by the inertio-capillary scales, $V_\gamma = \sqrt{\gamma/\rho_dD_0}$ and $\tau_\gamma = \sqrt{\rho_dD_0^3/\gamma}$, respectively. Furthermore, the pressure is normalized using the capillary pressure scale $p_\gamma = \gamma/D_0$. In such a conceptualization, $Oh$ and $We$ described in \S\ref{sec:intro} uniquely determine the system. 
The Ohnesorge number based on air viscosity $Oh_a = \left(\eta_a/\eta_d\right)Oh$ and air-drop density ratio $\rho_a/\rho_d$ are fixed at $10^{-5}$ and $10^{-3}$, respectively to minimize the influence of the surrounding medium on the impact forces. 
\Rev{Lastly, we keep the Bond number ($Bo$, see~\eqref{Eq:BoDefinition}) fixed at 1 throughout the manuscript. In our system, the relevance of gravity is characterized by the dimensionless Froude number $Fr = V_0^2/gD_0 = We/Bo$ comparing inertia with gravity. Throughout this manuscript, $Fr > 1$ and gravity's role is sub-dominant compared to inertia (for detailed discussion, see appendix~\ref{app:gravity}).}
The substrate is modeled as a no-slip and non-penetrable wall, whereas vanishing stress and pressure are applied at the remaining boundaries to mimic outflow conditions for the surrounding air. The domain boundaries are far enough from the drop not to influence its impact process ($\mathcal{L}_\text{max} \gg D_0$, $\mathcal{L}_\text{max} = 8R$ in the worst case). 
At $t = 0$, in our simulations, we release a spherical drop whose south pole is $0.05D_0$ away from the substrate and is falling with a velocity $V_0$. 
\Rev{It is important to note that experimental drops may deviate from perfect sphericity due to air drag post-needle release and potential residual oscillations from detachment. These shape perturbations are more pronounced in cases with low Weber and Ohnesorge numbers. To quantify this non-sphericity, we measure the drop's aspect ratio (horizontal to vertical diameter) immediately before substrate contact. The precise pre-impact drop shape can significantly influence subsequent impact dynamics \citep{thoraval-2013-jfm, yun2017bouncing,Zhang2019}. In our experiments, we constrain our analysis to drops with aspect ratios between 0.96 and 1.05. Given this narrow range, we posit that the impact of these shape variations is negligible compared to the experimental error bars derived from repeated trials under identical nominal conditions.}
The simulations utilize adaptive mesh refinement to finely resolve the velocity, viscous dissipation, and the VoF tracer fields. A minimum grid size $\Delta = D_0/2048$ is used for this study.

@@
